Near Optimal Behavior with Approximate State Abstraction

### 2016-02-04 paper/theory_results.tex
Unified notation for all theory results yay

%Model Lemma
\begin{lma}
\label{lma:model}
When $\mcS_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\forall_{s \in \mcS_G} : | V_G(s) - V_G^{\pi_{GA}}(s) | \leq \frac{2\epsilon + 2\gamma((|\mcS_G|-1)\epsilon)}{(1-\gamma)^3}
\end{equation}
\end{lma}

@@

Since the Q values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\epsilon = B$. Applying Lemma \ref{lma:Q*} yields Lemma \ref{lma:model}.

@@

Here we introduce the approximate abstraction family, $\phi_{bolt}$, which aggregates states with similar Boltzmann distributions on optimal Q values as equivalent. This family of abstractions is appealing as the Boltzmman distribution over Q-values has been of interest in the reinforcement learning literature due to its natural balancing of exploration and exploitation~\cite{sutton1998reinforcement}.

@@

We also assume that the difference in normalizing terms is bounded by some non-negative constant, $k\in \mathbb{R}$, of $\epsilon$:
\begin{equation}
\sum_b e^{Q^*(s_1,b)} - \sum_b e^{Q^*(s_2,b)} \leq k*\epsilon
\label{eq:bolt_denom}
\end{equation}

\begin{lma} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some non-negative constant $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in \mcS_G} : | V^{\pi^*_G}(s) - V^{\pi_{GA}}(s) | \leq \frac{2\epsilon\left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k\right)}{(1-\gamma)^2}
\end{equation}
\label{lma:bolt_lemma}
\end{lma}

We prove Lemma~\ref{lma:bolt_lemma} by using the approximation for $e^x$:
\begin{equation}
1 + x + \delta = e^x
\label{eq:e_to_x_approx}
\end{equation}
Where $\delta$ is the error associated with the approximation.

@@

By the approximation in Equation~\ref{eq:e_to_x_approx} and the assumption in Equation~\ref{eq:bolt_denom}:
\begin{align*}
|\frac{1 + Q(s_1,a) + \delta_1}{\sum_j e^{Q(s_1,a_j)}} - \frac{1 + Q(s_2,a) + \delta_2}{\sum_j e^{Q(s_1,a_j)} \pm k\epsilon}| \leq \epsilon
\end{align*}

It follows that:
\begin{equation}
|Q(s_1,a) - Q(s_2,a)| \leq \epsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k \right)
\label{eq:bolt_qs}
\end{equation}

Consequently, we can consider $\widetilde{\phi}_{bolt}$ as a type of $\epQ$, where the difference in Q values is bounded by: $\epsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k \right)$. The Lemma follows directly from Lemma~\ref{lma:Q*}.

@@

We consider a multinomial distribution over $Q^*$ for similar reasons as the Boltzmann.

\bdefn{$\ep_{mult}$}
We let $\phi_{mult}$ define a type of abstraction that, for fixed $\epsilon$ satisfies.
\begin{multline}
\ep_{mult}(s_1) = \ep_{mult}(s_2) \rightarrow \\
\forall_{a} \left|\frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)} - \frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)}\right| \leq \epsilon
\end{multline}
\edefn

We also assume that the difference in normalizing terms is bounded by some positive factor, $k \in \mathbb{R}$, of $\epsilon$:
\begin{equation}
|\sum_i Q(s_1,a_i) - \sum_j Q(s_2,a_j) | \leq k\epsilon
\end{equation}

\begin{lma} When $S_A$ is created using a function of the $\ep_{mult}$ type, for some non-negative constant $k \in \mathbb{R}$:
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{\frac{2\epsilon|\mathcal{A}|}{1-\gamma} + k \epsilon^2 + k}{(1-\gamma)^2}
\end{equation}
\end{lma}

@@

The proof follows a near identical strategy to that of Lemma~\ref{lma:bolt_lemma}.

>>>>

%Model Lemma
\begin{lma}
\label{lma:model}
When $\mcS_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\forall_{s \in \mcS_G} : | V_G^{\pi_G^*}(s) - V_G^{\pi_{GA}}(s) | \leq \frac{2\epsilon + 2\gamma((|\mcS_G|-1)\epsilon)}{(1-\gamma)^3}
\end{equation}
\end{lma}

@@

Since the $Q$-values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\epsilon = B$. Applying Lemma \ref{lma:Q*} yields Lemma \ref{lma:model}.

@@

Here we introduce the approximate abstraction family, $\phi_{bolt}$, which aggregates states with similar Boltzmann distributions on optimal $Q$-values as equivalent. This family of abstractions is appealing as the Boltzmman distribution over $Q$-values has been of interest in the reinforcement learning literature due to its natural balancing of exploration and exploitation~\cite{sutton1998reinforcement}.

@@

We also assume that the difference in normalizing terms is bounded by some non-negative constant, $k\in \mathbb{R}$, of $\epsilon$:
\begin{equation}
\left| \sum_b e^{Q_G^*(s_1,b)} - \sum_b e^{Q_G^*(s_2,b)} \right| \leq k*\epsilon
\label{eq:bolt_denom}
\end{equation}

\begin{lma} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some non-negative constant $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in \mcS_G} : | V_G^{\pi^*_G}(s) - V_G^{\pi_{GA}}(s) | \leq \frac{2\epsilon\left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k\right)}{(1-\gamma)^2}
\end{equation}
\label{lma:bolt_lemma}
\end{lma}

We prove Lemma~\ref{lma:bolt_lemma} by using the approximation for $e^x$:
\begin{equation}
 e^x = 1 + x + \delta  \approx 1 + x
\label{eq:e_to_x_approx}
\end{equation}
Where $\delta$ is the error associated with the approximation.

@@

By the approximation in Equation~\ref{eq:e_to_x_approx} and the assumption in Equation~\ref{eq:bolt_denom}:
\begin{align*}
|\frac{1 + Q_G(s_1,a) + \delta_1}{\sum_j e^{Q_G(s_1,a_j)}} - \frac{1 + Q_G(s_2,a) + \delta_2}{\sum_j e^{Q_G(s_1,a_j)} \pm k\epsilon}| \leq \epsilon
\end{align*}

It follows that:
\begin{equation}
|Q_G(s_1,a) - Q_G(s_2,a)| \leq \epsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k \right)
\label{eq:bolt_qs}
\end{equation}

Consequently, we can consider $\widetilde{\phi}_{bolt}$ as a type of $\epQ$, where the difference in Q values is bounded by: $\epsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k \right)$. Lemma \ref{lma:bolt_lemma} follows directly from Lemma \ref{lma:Q*}.

@@

We consider approximate abstractions derived from a multinomial distribution over $Q^*$ for similar reasons to our motivation in considering the Boltzmann distribution.

\bdefn{$\ep_{mult}$}
We let $\phi_{mult}$ define a type of abstraction that, for fixed $\epsilon$ satisfies.
\begin{multline}
\ep_{mult}(s_1) = \ep_{mult}(s_2) \rightarrow \\
\forall_{a} \left|\frac{Q_G^*(s_1,a)}{\sum_b Q_G^*(s_1,b)} - \frac{Q_G^*(s_1,a)}{\sum_b Q_G^*(s_1,b)}\right| \leq \epsilon
\end{multline}
\edefn

We also assume that the difference in normalizing terms is bounded by some positive factor, $k \in \mathbb{R}$, of $\epsilon$:
\begin{equation}
\left |\sum_i Q_G(s_1,a_i) - \sum_j Q_G(s_2,a_j) \right | \leq k\epsilon
\end{equation}

\begin{lma} When $S_A$ is created using a function of the $\ep_{mult}$ type, for some non-negative constant $k \in \mathbb{R}$:
\begin{equation}
\forall_{s \in S_M} : | V_G^{\pi^*_G}(s) - V_G^{\pi_{GA}}(s) | \leq \frac{\frac{2\epsilon|\mathcal{A}|}{1-\gamma} + k \epsilon^2 + k}{(1-\gamma)^2}
\end{equation}
\end{lma}

@@

The proof follows an identical strategy to that of Lemma~\ref{lma:bolt_lemma} but without the use of the approximation for $e^x$.
